D3064R0: OOTA Execution on Real Implementations is Provably Vacuous

### 2023-12-16 ootavacuous.tex
Add if-then-else example to per-execution sdep section
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>
From sdep U rf to sdep U rfe plus minor edits
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

Others, current authors included, have defined an OOTA cycle as a
cycle in sdep $\cup$ rf, where sdep is the set of semantic
dependencies within each thread and rf is the set of store-to-load
links, whether intra-thread or between threads.
This is a fine definition, and is consistent with the words in the C++
standard, except that it simply shunts all the complexity onto the
term ``semantic dependency''.

@@

@@

\subsubsection{Implementers and Users Influence ``Semantic Dependencies'' Definition}
\label{sec:Implementers and Users Influence ``Semantic Dependencies'' Definition}

@@

@@

Third, one could imagine implementations whose analysis is restricted to
a thread.
In this case, the analysis stops at loads to objects that were last
stored to by other threads (``rfe''), stores to objects where the next
and/or previous stores in that object's modification order were executed
by other threads (``coe'')), and loads from objects for which the value
returned by a given load was overwritten by some other thread
(``fre'').\footnote{
	See Appendix~\ref{app:Inter-Process Communications} for more
	information on these memory-reference interprocess communications
	modes.}

>>>>

Others have defined an OOTA cycle as a cycle in sdep $\cup$ rf, where
sdep is the set of semantic dependencies within each thread and rf is
the set of store-to-load links, whether intra-thread (rfi) or between
threads (rfe).\footnote{
	See Appendix~\ref{app:Inter-Process Communications} for definitions
	and properies of rf, rfe, and rfi.}
This is a fine definition, and is consistent with the words in the C++
standard, except that it simply shunts all the complexity onto the
term ``semantic dependency''.

The current authors deviate slightly from this tradition by instead
defining an OOTA cycle as a cycle in sdep $\cup$ rfe, where rfe is the
set of store-to-load links, but excluding intra-thread rfi links, which
are instead considered to be part of sdep.
Although this does shunt even more complexity onto the term
``semantic dependency'', it also enables us to bring to bear
laws of physics and properties of hardware onto correct real-world
C++ implementations.\footnote{
	Thanks to Alan Stern for providing a litmus test that demonstrated
	that we had unwittingly (but productively!) shifted our definition
	to sdep $\cup$ rfe.}

@@

Finally, consider this:

\begin{quote}
	\co{r1 = x;} \\
	\co{if (r1 > 0)} \\
	\co{\ \ \ \ y = r1;} \\
	\co{else} \\
	\co{\ \ \ \ z = r1;}
\end{quote}

Here there is a semantic dependency from \co{x}, but in some executions
it extends to \co{y} and in others to \co{z}.

@@

\subsubsection{Implementers and Users Influence Semantic-Dependencies Definition}
\label{sec:Implementers and Users Influence Semantic-Dependencies Definition}

@@

\emph{A definition of sdep drawn strictly from the standard will be
at best an approximation to a definition that is useful in practice.}

@@

Third, one could imagine implementations whose analysis is restricted to
a thread.
In this case, the analysis stops at loads to objects that were last
stored to by other threads (``rfe''), stores to objects where the next
and/or previous stores in that object's modification order were executed
by other threads (``coe'')), and loads from objects for which the value
returned by a given load was overwritten by some other thread
(``fre'').\footnote{
	See Appendix~\ref{app:Inter-Process Communications} for more
	information on these memory-reference interprocess communications
	links.}
